---
title: "Child Poverty Talking Points"
title-block-banner: TRUE
fig-dpi: 300
fig-width: 6
fig-height: 4
warning: FALSE
error: TRUE
echo: FALSE
message: FALSE
---

```{r, Load Data}
#| include: FALSE
#| 
#####  Load Libraries #####
library(tidycensus)
library(tidyverse)
library(rio)
library(sf)
#library(directlabels)
library(cpaltemplates)
#library(terra)
#library(magick)
library(plotly)
library(DT)

#datalib <- "E:/CPAL Dropbox/" # Michael Desktop Directory
datalib <- "C:/Users/Michael/CPAL Dropbox/" # Michael Laptop Directory

##### Classify Objects and ACS Variables #####
yearsOld <- lst(2010, 2011)
yearsNew <- lst(2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2021, 2022)

acs_s <- load_variables(2018, "acs1/subject", cache = TRUE)
#acs_b <- load_variables(2011, "acs5", cache = TRUE)
#decennial <- load_variables(2020, "as", cache = TRUE)

acs_varOld <- c(
  tot_pop = "S1701_C01_001", #total population
  pop_u18 = "S1701_C01_002", #population under 18
  pop_bp = "S1701_C02_001", #population below poverty
  bp_u18 = "S1701_C02_002", #population under 18 below poverty
  
  mhi = "S0501_C01_101", #median household income
  
  his_pop = "S1701_C01_016", #hispanic non-white population
  blk_pop = "S1701_C01_010", #black population population
  asn_pop = "S1701_C01_012", #asian population population
  wht_pop = "S1701_C01_017", #white non-hispanic popuation
  
  his_bp = "S1701_C02_016", #hispanic non-white population below poverty
  blk_bp = "S1701_C02_010", #black population population below poverty
  asn_bp = "S1701_C02_012", #asian population population below poverty
  wht_bp = "S1701_C02_017" #white non-hispanic popuation below poverty
)

acs_varNew <- c(
  tot_pop = "S1701_C01_001", #total population
  pop_u18 = "S1701_C01_002", #population under 18
  pop_u5 = "S1701_C01_003", #population under 5
  pop_bp = "S1701_C02_001", #population below poverty
  bp_u18 = "S1701_C02_002", #population under 18 below poverty
  bp_u5 = "S1701_C02_003", #population under 5 below poverty
  
  mhi = "S0501_C01_101", #median household income
  
  his_pop = "S1701_C01_020", #hispanic non-white population
  blk_pop = "S1701_C01_014", #black population population
  asn_pop = "S1701_C01_016", #asian population population
  wht_pop = "S1701_C01_021", #white non-hispanic popuation
  
  his_bp = "S1701_C02_020", #hispanic non-white population below poverty
  blk_bp = "S1701_C02_014", #black population population below poverty
  asn_bp = "S1701_C02_016", #asian population population below poverty
  wht_bp = "S1701_C02_021" #white non-hispanic popuation below poverty
)

##### Pull Data from 2010 to 2020 #####
cpal_old <- map(
  yearsOld,
  ~get_acs(
    geography = "place", 
    variables = acs_varOld,
    year = .x,
    survey = "acs1", 
    output = "wide"),
  ) %>%
  map2(yearsOld, ~mutate(.x, YEAR = .y)) %>%
  reduce(., rbind) %>%
  mutate(NAME = str_replace(NAME, " city", ""))

cpal_2020 <- get_acs(
    geography = "place", 
    variables = acs_varNew,
    year = 2020, 
    survey = "acs5", 
    output = "wide") %>%
  mutate(NAME = str_replace(NAME, " city", ""),
         YEAR = 2020)

cpal_multi <- map(
  yearsNew,
  ~get_acs(
    geography = "place", 
    variables = acs_varNew,
    year = .x, 
    survey = "acs1", 
    output = "wide"),
  ) %>%
  map2(yearsNew, ~mutate(.x, YEAR = .y)) %>%
  reduce(., rbind) %>%
  mutate(NAME = str_replace(NAME, " city", ""))

#calculate 3 year average using lag function and long format data
yr3Avg <- function(x){
  round((x+lag(x, n = 1)+lag(x, n = 3))/3, digits = 0)
}

#calculate year to year percent change using lag function and long format data
yrPerCh <- function(x){
  (x - lag(x, n = 1))/lag(x, n = 1)
}

##### Join into Complete Dataset #####
cpal_comp <- full_join(cpal_multi, cpal_2020) %>%
  full_join(., cpal_old) %>%
  rename(tot_pop_1y = tot_popE,
         pop_u18_1y = pop_u18E,
         pop_u5_1y = pop_u5E,
         pop_bp_1y = pop_bpE,
         bp_u18_1y = bp_u18E,
         bp_u5_1y = bp_u5E,
         mhi = mhiE) %>%
  select(-ends_with("M")) %>%
  relocate(GEOID, NAME, YEAR, everything(.)) %>%
  group_by(NAME) %>%
  arrange(NAME, YEAR) %>%
  mutate(tot_pop_3y = yr3Avg(x = tot_pop_1y),
         pop_u18_3y = yr3Avg(x = pop_u18_1y),
         pop_u5_3y = yr3Avg(x = pop_u5_1y),
         pop_bp_3y = yr3Avg(x = pop_bp_1y),
         bp_u18_3y = yr3Avg(x = bp_u18_1y),
         bp_u5_3y = yr3Avg(x = bp_u5_1y),
         
         tot_pop_ch = yrPerCh(x = tot_pop_1y),
         pop_u18_ch = yrPerCh(x = pop_u18_1y),
         pop_u5_ch = yrPerCh(x = pop_u5_1y),
         pop_bp_ch = yrPerCh(x = pop_bp_1y),
         bp_u18_ch = yrPerCh(x = bp_u18_1y),
         bp_u5_ch = yrPerCh(x = bp_u5_1y),
         
         c5bp_1y = bp_u5_1y/pop_u5_1y,
         c5bp_3y = bp_u5_3y/pop_u5_3y,
         cbp_1y = bp_u18_1y/pop_u18_1y,
         cbp_3y = bp_u18_3y/pop_u18_3y,
         bp_1y = pop_bp_1y/tot_pop_1y,
         bp_3y = pop_bp_3y/tot_pop_3y,
         
         oth_bpE = pop_bp_1y-(his_bpE+asn_bpE+wht_bpE+blk_bpE),
         oth_popE = tot_pop_1y-(his_popE+asn_popE+wht_popE+blk_popE),
         
         oth_bp_per = oth_bpE/oth_popE,
         his_bp_per = his_bpE/his_popE,
         asn_bp_per = asn_bpE/asn_popE,
         blk_bp_per = blk_bpE/blk_popE,
         wht_bp_per = wht_bpE/wht_popE
         ) %>%
  ungroup() %>%
  mutate(highlight=ifelse(NAME == "Dallas, Texas", "DAL", "NONE"),
         highlight = factor(highlight, levels = c("NONE", "DAL")))

#names(cpal_comp)

##### Filter out multiple groups based on population size #####
cpal_1mil <- cpal_comp %>%
  group_by(NAME) %>%
  filter(any(tot_pop_1y >= 1000000)) %>%
  mutate(CODE = ifelse(NAME == "Chicago, Illinois", "CHI", 
                       ifelse(NAME == "Dallas, Texas", "DAL", 
                              ifelse(NAME == "Houston, Texas", "HOU", 
                                     ifelse(NAME == "Los Angeles, California", "LA", 
                                            ifelse(NAME == "New York, New York", "NYC", 
                                                   ifelse(NAME == "Philadelphia, Pennsylvania", "PHI", 
                                                          ifelse(NAME == "Phoenix, Arizona", "PHO", 
                                                                 ifelse(NAME == "San Antonio, Texas", "SAN", 
                                                                        ifelse(NAME == "San Diego, California", "SD", 
                                                                               ifelse(NAME == "San Jose, California", "SJ", "ERROR")))))))))))

cpal_1milCh <- cpal_1mil %>%
  filter(YEAR == 2017 | YEAR == max(cpal_comp$YEAR)) %>%
  select(GEOID, NAME, CODE, YEAR, highlight, cbp_3y, cbp_1y) %>%
  pivot_wider(names_from = YEAR, values_from = c(cbp_3y, cbp_1y)) %>%
  rename(OLD3y = 5,
         NEW3y = 6,
         OLD1y = 7,
         NEW1y = 8,
         ) %>%
  mutate(cbpChange1y = (NEW1y-OLD1y),
         cbpChange3y = (NEW3y-OLD3y))

cpal_1mil1YrCh <- cpal_1mil %>%
  filter(YEAR == 2021 | YEAR == max(cpal_comp$YEAR)) %>%
  select(GEOID, NAME, CODE, YEAR, highlight, cbp_3y, cbp_1y) %>%
  pivot_wider(names_from = YEAR, values_from = c(cbp_3y, cbp_1y)) %>%
  rename(OLD3y = 5,
         NEW3y = 6,
         OLD1y = 7,
         NEW1y = 8,
         ) %>%
  mutate(cbpChange1y = (NEW1y-OLD1y),
         cbpChange3y = (NEW3y-OLD3y))

# dallas, ft worth, houston, san antonio, austin, el paso
cpal_texas <- cpal_comp %>%
  filter(str_detect(NAME, ", Texas")) %>%
  group_by(NAME) %>%
  filter(any(tot_pop_1y >= 500000)) %>%
  mutate(CODE = ifelse(NAME == "Dallas, Texas", "DAL", 
                              ifelse(NAME == "Houston, Texas", "HOU", 
                                     ifelse(NAME == "San Antonio, Texas", "SAN", 
                                            ifelse(NAME == "El Paso, Texas", "ELP", 
                                                   ifelse(NAME == "Austin, Texas", "AUS", 
                                                          ifelse(NAME == "Fort Worth, Texas", "FTW", "ERROR"))))))) %>%
  filter(CODE != "FTW" & CODE != "AUS")

cpal_texasCh <- cpal_texas %>%
  filter(YEAR == 2017 | YEAR == max(cpal_comp$YEAR)) %>%
  select(GEOID, NAME, CODE, YEAR, highlight, cbp_3y, cbp_1y) %>%
  pivot_wider(names_from = YEAR, values_from = c(cbp_3y, cbp_1y)) %>%
  rename(OLD3y = 5,
         NEW3y = 6,
         OLD1y = 7,
         NEW1y = 8,
         ) %>%
  mutate(cbpChange1y = (NEW1y-OLD1y),
         cbpChange3y = (NEW3y-OLD3y))

cpal_texas1YrCh <- cpal_texas %>%
  filter(YEAR == 2021 | YEAR == max(cpal_comp$YEAR)) %>%
  select(GEOID, NAME, CODE, YEAR, highlight, cbp_3y, cbp_1y) %>%
  pivot_wider(names_from = YEAR, values_from = c(cbp_3y, cbp_1y)) %>%
  rename(OLD3y = 5,
         NEW3y = 6,
         OLD1y = 7,
         NEW1y = 8,
         ) %>%
  mutate(cbpChange1y = (NEW1y-OLD1y),
         cbpChange3y = (NEW3y-OLD3y))

cpal_dallas <- cpal_texas %>%
  filter(CODE == "DAL")

##### Data Table #####
tbl_1mil <- cpal_1mil %>%
  filter(YEAR %in% c(2022, 2021, 2017)) %>%
  select(NAME, YEAR, pop_u18_1y, bp_u18_1y, cbp_1y, pop_u18_3y, bp_u18_3y, cbp_3y) %>%
  pivot_wider(names_from = YEAR, values_from = c(pop_u18_1y:cbp_3y))
         
rm(cpal_2020)
rm(cpal_multi)
rm(cpal_old)

rm(yearsNew)
rm(yearsOld)

# pasted text poverty metrics

yearMax <- max(cpal_dallas$YEAR)
dallasMax <- filter(cpal_dallas, YEAR == yearMax)
dallasMin <- filter(cpal_dallas, YEAR == 2017)

cbpMax <- scales::percent(dallasMax$cbp_3y, accuracy = 0.1)
bp_u18Max <- dallasMax$bp_u18_3y
bp_u18Min <- dallasMin$bp_u18_3y

bp_u18Diff <- scales::number(bp_u18Max-bp_u18Min, big.mark = ",")

bp_u18Chg <- scales::percent((bp_u18Max-bp_u18Min)/bp_u18Min, accuracy = 0.1)

bp_u18Max <- scales::number(dallasMax$bp_u18_3y, big.mark = ",")
```

# Child Poverty in the City of Dallas

The Child Poverty Rate in the City of Dallas in `r yearMax` is `r cbpMax`

According to estimates from the American Community Survey for `r yearMax`, there are now `r paste0(bp_u18Diff, " ", ifelse(bp_u18Diff > 0, "greater", "fewer"))` children in poverty within the City of Dallas than there were in 2017.

This poverty rate for all data included in this report is calculated from the 3 year average of 1-year American Community Survey Estimates unless specified as otherwise.

```{r, Child Poverty in Dallas}
plotcbpLine <- cpal_dallas %>%
  filter(YEAR >= 2017) %>%
  select(NAME, YEAR, CODE, cbp_3y, cbp_1y) %>%
  pivot_longer(cols = c(cbp_3y, cbp_1y)) %>%
  mutate(name = ifelse(name == "cbp_3y", "ACS 3-Year Average", "ACS 1-Year Survey")) %>%
ggplot(aes(x=YEAR, y=value, color = name)) +
  geom_line(linewidth = 1, alpha = 0.8) +
#  geom_text(aes(label = paste0(round(value*100, digits = 1), "%")),
#            position = position_stack(vjust = 1.1),
#            color = palette_cpal_main[3],
#            fontface = "bold") +
  geom_point(size = 3) +
  scale_color_manual(values = palette_cpal_main) +
  scale_size_manual(values=c(2, 1)) +
  scale_y_continuous(limits = c(0, 0.40),
                     breaks = c(0, 0.1, 0.2, 0.3, 0.4),
                     labels = scales::percent,
                     expand = c(0, 0)) +
  scale_x_continuous(limits = c(2016.8, 2022.2),
                     breaks = c(2017, 2018, 2019, 2020, 2021, 2022),
                     expand = c(0.005, 0.005)
  ) +
  labs(x = "",
       y = "") +
  theme_cpal_print() #+
#  theme(legend.position = "NONE")

cpal_plot(zeroGrob(),
          cpal_title(string = "Child Poverty Rate in the City of Dallas",
                     width = 75),
          plotcbpLine, 
          cpal_source(text = "U.S. Census Bureau. American Community Survey Table S1702, 1-year estimates. (2017-2022)"),
          cpal_note(text = "2020 ACS Estimates are derived from 5-Year ACS due to pandemic collection issues."),
          zeroGrob(),
          cpal_logo_text(), 
          ncol = 1,
          heights = c(1, 3, 30, 1, 1, 1, 2))

```

### Child Poverty Metrics Table

```{r, Child Poverty in Dallas Table}
cpal_dallas %>%
  ungroup(.) %>%
  filter(YEAR >= 2017) %>%
  select(NAME, YEAR, CODE, cbp_3y, cbp_1y, bp_u18_1y, bp_u18_3y) %>%
  mutate(childpov1y = paste0(scales::number(bp_u18_1y, big.mark = ","), " (", scales::percent(cbp_1y, accuracy = 0.1), ")"),
         childpov3y = paste0(scales::number(bp_u18_3y, big.mark = ","), " (", scales::percent(cbp_3y, accuracy = 0.1), ")")) %>%
  select(-NAME, -CODE, -(cbp_3y:bp_u18_3y)) %>%
  pivot_longer(cols = c(childpov3y, childpov1y)) %>%
  mutate(name = ifelse(name == "childpov3y", "ACS 3-Year Average", "ACS 1-Year Survey")) %>%
  pivot_wider(names_from = YEAR, values_from = value) %>%
  rename('Data Source' = name) %>%
  datatable(options = list(paging = FALSE,
                           searching = FALSE,
                           info = FALSE
                           ),
            rownames = FALSE) %>%
  formatStyle(columns = colnames(.), fontWeight = 'bold', `text-align` = 'left') %>%
  formatStyle(columns = colnames(.), fontFamily = "Poppins")
```

```{r, Poverty by Race/Ethnicity}
plotRaceLine <- cpal_dallas %>%
  select(GEOID, NAME, YEAR, bp_1y, his_bp_per:wht_bp_per) %>%
  pivot_longer(cols = bp_1y:wht_bp_per, names_to = "race", values_to = "percent") %>%
  mutate(race = ifelse(race == "his_bp_per", "Hispanic/Latino",
                       ifelse(race == "blk_bp_per", "Black or African American",
                              ifelse(race == "asn_bp_per", "Asian or Pacific Islander",
                                     ifelse(race == "wht_bp_per", "White Non-Hispanic", 
                                            ifelse(race == "bp_1y", "Overall Population", "ERROR"))))),
         race = factor(race, levels = c("Overall Population", "White Non-Hispanic", "Hispanic/Latino", "Black or African American", "Asian or Pacific Islander", "ERROR"))) %>%
  filter(YEAR >= 2017) %>%
  ggplot(aes(x=YEAR, y=percent, color = race, group = race)) +
  geom_line(linewidth = 1, alpha = 0.8) +
  geom_point(size = 3) +
  scale_color_manual(values = palette_cpal_main) +
  scale_size_manual(values=c(2, 1)) +
  scale_y_continuous(limits = c(0, 0.35),
                     breaks = c(0, 0.1, 0.2, 0.3),
                     labels = scales::percent,
                     expand = c(0, 0)) +
  scale_x_continuous(limits = c(2016.8, 2022.2),
                     breaks = c(2017, 2018, 2019, 2020, 2021, 2022),
                     expand = c(0.005, 0.005)
  ) +
  labs(x = "",
       y = "") +
  theme_cpal_print() +
  guides(color = guide_legend(nrow = 2))

cpal_plot(zeroGrob(),
          cpal_title(string = "Poverty Rate by Race/Ethnicity in the City of Dallas",
                     width = 75),
          plotRaceLine, 
          cpal_source(text = "U.S. Census Bureau. American Community Survey Table S1702, 1-year estimates. (2017-2022)"),
          cpal_note(text = "2020 ACS Estimates are derived from 5-Year ACS due to pandemic collection issues."),
          zeroGrob(),
          cpal_logo_text(), 
          ncol = 1,
          heights = c(1, 3, 30, 1, 1, 1, 2))

```

# Child Poverty in Peer Metropolitan Cities

Compared to cities with a population over 1,000,000, the City Dallas has the fourth highest Child Poverty Rate in `r yearMax`

```{r, Child Poverty Rate in Peer Metropolitan Cities}
plot1MilLine <- cpal_1mil %>%
  mutate(CODE = factor(CODE, levels = c("DAL", "NYC", "LA", "CHI", "HOU", "PHO", "SAN", "PHI", "SD", "SJ"))) %>%
  filter(YEAR >= 2017) %>%
  ggplot(aes(x=YEAR, y=cbp_3y, group = CODE, color = highlight)) +
  geom_line(linewidth = 1.5) +
  geom_point(size = 2) +
  scale_color_manual(values = palette_cpal_main) +
  scale_y_continuous(limits = c(0, 0.45),
                     breaks = c(0, 0.1, 0.2, 0.3, 0.4),
                     labels = scales::percent,
                     expand = c(0, 0)) +
  scale_x_continuous(limits = c(2016.8, 2022.2),
                     breaks = c(2017, 2018, 2019, 2020, 2021, 2022),
                     expand = c(0.005, 0.005)
  ) +
  labs(x = "",
       y = "") +
  theme_cpal_print() +
  facet_wrap(~CODE, ncol = 5) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  theme(legend.position = "NONE")

cpal_plot(zeroGrob(),
          cpal_title(string = "Child Poverty Rate in Metropolitan Cities",
                     width = 75),
          plot1MilLine, 
          cpal_source(text = "U.S. Census Bureau. American Community Survey Table S1702, 1-year estimates. (2017-2022)"),
          cpal_note(text = "2020 ACS Estimates are derived from 5-Year ACS due to pandemic collection issues."),
          zeroGrob(),
          cpal_logo_text(), 
          ncol = 1,
          heights = c(1, 3, 30, 1, 1, 1, 2))


```

Between 2021 and 2022 the 3-year average shows that the Child Poverty RAte in all large cities continues to decrease relative to prior years.

```{r, One Year Child Poverty Change in Cities Over 1,000,000 Population}
plot1MilBar <- cpal_1mil1YrCh %>%
  ungroup() %>%
  select(CODE, highlight, cbpChange1y, cbpChange3y) %>%
  pivot_longer(cols = c(cbpChange3y, cbpChange1y)) %>%
  mutate(name = ifelse(name == "cbpChange3y", "ACS 3-Year Average", "ACS 1-Year Survey")) %>%
  mutate(CODE = factor(CODE, levels = c("DAL", "NYC", "LA", "CHI", "HOU", "PHO", "SAN", "PHI", "SD", "SJ"))) %>%
  ggplot(aes(x=value, y=CODE, fill = highlight)) +
  geom_bar(stat = "identity", width = 0.8) +
  geom_text(aes(label = paste0(round(value*100, digits = 1), "%"),
                x = 0.035),
                color = palette_cpal_main[3],
                fontface = "bold") +
  facet_wrap(~name) +
  scale_fill_manual(values = palette_cpal_main) +
  scale_x_continuous(labels = scales::percent,
                     expand = c(0, 0)) +
  labs(x = "",
       y = "") +
  theme_cpal_print() +
  theme(legend.position = "NONE")

cpal_plot(zeroGrob(),
          cpal_title(string = paste0("Percent Change in Child Poverty in Metropolitan Cities (2021 to ", max(cpal_1mil$YEAR), ")"),
                     width = 75),
          plot1MilBar, 
          cpal_source(text = "U.S. Census Bureau. American Community Survey Table S1702, 1-year estimates. (2017-2022)"),
          cpal_note(text = "2020 ACS Estimates are derived from 5-Year ACS due to pandemic collection issues."),
          zeroGrob(),
          cpal_logo_text(), 
          ncol = 1,
          heights = c(1, 3, 30, 1, 1, 1, 2))


```

### Child Poverty Metrics in Peer Metropolitan Cities Table
```{r, Child Poverty Table Peer Metropoloitan Cities}
cpal_1mil %>%
  ungroup(.) %>%
  filter(YEAR %in% c(2022)) %>%
  select(NAME, YEAR, CODE, cbp_3y, cbp_1y, bp_u18_1y, bp_u18_3y) %>%
  mutate(childpov1y = paste0(scales::number(bp_u18_1y, big.mark = ","), " (", scales::percent(cbp_1y, accuracy = 0.1), ")"),
         childpov3y = paste0(scales::number(bp_u18_3y, big.mark = ","), " (", scales::percent(cbp_3y, accuracy = 0.1), ")")) %>%
  select(-YEAR, -NAME, -(cbp_3y:bp_u18_3y)) %>%
  pivot_longer(cols = c(childpov3y, childpov1y)) %>%
  pivot_wider(names_from = name, values_from = value) %>%
  rename('City' = CODE,
         'ACS 3-Year Average' = childpov3y,
         'ACS 1-Year Survey' = childpov1y) %>%
  datatable(options = list(paging = FALSE,
                           searching = FALSE,
                           info = FALSE
                           ),
            rownames = FALSE) %>%
  formatStyle(columns = colnames(.), fontWeight = 'bold', `text-align` = 'left') %>%
  formatStyle(columns = colnames(.), fontFamily = "Poppins")
```

Since 2017 the City of Dallas has had the fifth largest drop in Child Poverty compared to its peer metropolitan cities in the 3-year average.

```{r, Child Poverty Change in Peer Metropolitan Cities}
plot1MilBar <- cpal_1milCh %>%
  ungroup() %>%
  select(CODE, highlight, cbpChange1y, cbpChange3y) %>%
  pivot_longer(cols = c(cbpChange3y, cbpChange1y)) %>%
  mutate(name = ifelse(name == "cbpChange3y", "ACS 3-Year Average", "ACS 1-Year Survey")) %>%
  mutate(CODE = factor(CODE, levels = c("DAL", "NYC", "LA", "CHI", "HOU", "PHO", "SAN", "PHI", "SD", "SJ"))) %>%
  ggplot(aes(x=value, y=CODE, fill = highlight)) +
  geom_bar(stat = "identity", width = 0.8) +
  geom_text(aes(label = paste0(round(value*100, digits = 1), "%"),
                x = 0.008),
                color = palette_cpal_main[3],
                fontface = "bold") +
  facet_wrap(~name) +
  scale_fill_manual(values = palette_cpal_main) +
  scale_x_continuous(labels = scales::percent,
                     expand = c(0, 0)) +
  labs(x = "",
       y = "") +
  theme_cpal_print() +
  theme(legend.position = "NONE")

cpal_plot(zeroGrob(),
          cpal_title(string = paste0("Percent Change in Child Poverty in Metropolitan Cities (2017 to ", max(cpal_1mil$YEAR), ")"),
                     width = 75),
          plot1MilBar, 
          cpal_source(text = "U.S. Census Bureau. American Community Survey Table S1702, 1-year estimates. (2017-2022)"),
          cpal_note(text = "2020 ACS Estimates are derived from 5-Year ACS due to pandemic collection issues."),
          zeroGrob(),
          cpal_logo_text(), 
          ncol = 1,
          heights = c(1, 3, 30, 1, 1, 1, 2))


```

```{r, Poverty by Race/Ethnicity in Peer Metropolitan Cities}
plotRace1Mil <- cpal_1mil %>%
  select(GEOID, NAME, CODE, YEAR, bp_1y, his_bp_per:wht_bp_per) %>%
  pivot_longer(cols = bp_1y:wht_bp_per, names_to = "race", values_to = "percent") %>%
  mutate(race = ifelse(race == "his_bp_per", "Hispanic/Latino",
                       ifelse(race == "blk_bp_per", "Black or African American",
                              ifelse(race == "asn_bp_per", "Asian or Pacific Islander",
                                     ifelse(race == "wht_bp_per", "White Non-Hispanic", 
                                            ifelse(race == "bp_1y", "Overall Population", "ERROR"))))),
         race = factor(race, levels = c("Overall Population", "White Non-Hispanic", "Hispanic/Latino", "Black or African American", "Asian or Pacific Islander", "ERROR"))) %>%
  filter(YEAR >= max(cpal_1mil$YEAR)) %>%
  ggplot(aes(x=percent, y=CODE, color = race)) +
  geom_point(size = 2) +
  scale_color_manual(values = palette_cpal_main) +
  scale_x_continuous(limits = c(0, 0.32),
                     breaks = c(0, 0.1, 0.2, 0.3),
                     labels = scales::percent,
                     expand = c(0, 0)) +
  labs(x = "",
       y = "") +
  theme_cpal_print() +
  theme(legend.direction = "vertical",
        legend.position = "right")

cpal_plot(zeroGrob(),
          cpal_title(string = "Poverty Rate by Race/Ethnicity in Metropolitan Cities",
                     width = 75),
          plotRace1Mil, 
          cpal_source(text = "U.S. Census Bureau. American Community Survey Table S1702, 1-year estimates. (2017-2022)"),
          cpal_note(text = "2020 ACS Estimates are derived from 5-Year ACS due to pandemic collection issues."),
         zeroGrob(),
         cpal_logo_text(), 
          ncol = 1,
          heights = c(1, 3, 30, 1, 1, 1, 2))


```

# Child Poverty in Peer Texas Cities

Compared to large neighboring Texas cities, the City Dallas has the fourth highest Child Poverty Rate in `r yearMax` in Texas.

```{r, Child Poverty Rate in Texas Cities}
plot1MilLine <- cpal_texas %>%
  mutate(CODE = factor(CODE, levels = c("DAL", "HOU", "ELP", "AUS", "SAN", "FTW"))) %>%
  filter(YEAR >= 2017) %>%
  ggplot(aes(x=YEAR, y=cbp_3y, group = CODE, color = highlight)) +
  geom_line(linewidth = 1.5) +
  geom_point(size = 3) +
  scale_color_manual(values = palette_cpal_main) +
  scale_y_continuous(limits = c(0, 0.45),
                     breaks = c(0, 0.1, 0.2, 0.3, 0.4),
                     labels = scales::percent,
                     expand = c(0, 0)) +
  scale_x_continuous(limits = c(2016.8, 2022.2),
                     breaks = c(2017, 2018, 2019, 2020, 2021, 2022),
                     expand = c(0.005, 0.005)
  ) +
  labs(x = "",
       y = "") +
  theme_cpal_print() +
  facet_wrap(~CODE, ncol = 2) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  theme(legend.position = "NONE")

cpal_plot(zeroGrob(),
          cpal_title(string = "Child Poverty Rate in Peer Texas Cities",
                     width = 75),
          plot1MilLine, 
          cpal_source(text = "U.S. Census Bureau. American Community Survey Table S1702, 1-year estimates. (2017-2022)"),
          cpal_note(text = "2020 ACS Estimates are derived from 5-Year ACS due to pandemic collection issues."),
          zeroGrob(),
          cpal_logo_text(), 
          ncol = 1,
          heights = c(1, 3, 30, 1, 1, 1, 2))


```

Since 2017 the City of Dallas has had the `largest drop in Child Poverty compared to its peer Texas cities.

```{r, Child Poverty Change in Texas Cities}
plotTexasBar <- cpal_texasCh %>%
  ungroup() %>%
  select(CODE, highlight, cbpChange1y, cbpChange3y) %>%
  pivot_longer(cols = c(cbpChange3y, cbpChange1y)) %>%
  mutate(name = ifelse(name == "cbpChange3y", "ACS 3-Year Average", "ACS 1-Year Survey")) %>%
  mutate(CODE = factor(CODE, levels = c("DAL", "HOU", "ELP", "AUS", "SAN", "FTW"))) %>%
  ggplot(aes(x=value, y=CODE, fill = highlight)) +
  geom_bar(stat = "identity", width = 0.8) +
  geom_text(aes(label = paste0(round(value*100, digits = 1), "%"),
                x = 0.029),
                color = palette_cpal_main[3],
                fontface = "bold") +
  facet_wrap(~name) +
  scale_fill_manual(values = palette_cpal_main) +
  scale_x_continuous(labels = scales::percent,
                     expand = c(0, 0)) +
  labs(x = "",
       y = "") +
  theme_cpal_print() +
  theme(legend.position = "NONE")

cpal_plot(zeroGrob(),
          cpal_title(string = paste0("Percent Change in Child Poverty in Peer Texas Cities (2017 to ", max(cpal_1mil$YEAR), ")")),
          plotTexasBar, 
          cpal_source(text = "U.S. Census Bureau. American Community Survey Table S1702, 1-year estimates. (2017-2022)"),
          cpal_note(text = "2020 ACS Estimates are derived from 5-Year ACS due to pandemic collection issues."),
          zeroGrob(),
          cpal_logo_text(), 
          ncol = 1,
          heights = c(1, 3, 30, 1, 1, 1, 2))
```

```{r, Poverty by Race/Ethnicity in Texas Cities}
plotRaceTexas <- cpal_texas %>%
  select(GEOID, NAME, CODE, YEAR, bp_1y, his_bp_per:wht_bp_per) %>%
  pivot_longer(cols = bp_1y:wht_bp_per, names_to = "race", values_to = "percent") %>%
  mutate(race = ifelse(race == "his_bp_per", "Hispanic/Latino",
                       ifelse(race == "blk_bp_per", "Black or African American",
                              ifelse(race == "asn_bp_per", "Asian or Pacific Islander",
                                     ifelse(race == "wht_bp_per", "White Non-Hispanic", 
                                            ifelse(race == "bp_1y", "Overall Population", "ERROR"))))),
         race = factor(race, levels = c("Overall Population", "White Non-Hispanic", "Hispanic/Latino", "Black or African American", "Asian or Pacific Islander", "ERROR"))) %>%
  filter(YEAR >= max(cpal_1mil$YEAR)) %>%
  ggplot(aes(x=percent, y=CODE, color = race)) +
  geom_point(size = 2) +
  scale_color_manual(values = palette_cpal_main) +
  scale_x_continuous(limits = c(0, 0.32),
                     breaks = c(0, 0.1, 0.2, 0.3),
                     labels = scales::percent,
                     expand = c(0, 0)) +
  labs(x = "",
       y = "") +
  theme_cpal_print() +
  theme(legend.direction = "vertical",
        legend.position = "right")


cpal_plot(zeroGrob(),
          cpal_title(string = "Poverty Rate by Race/Ethnicity in Texas Cities",
                     width = 75),
          plotRaceTexas, 
          cpal_source(text = "U.S. Census Bureau. American Community Survey Table S1702, 1-year estimates. (2017-2022)"),
          cpal_note(text = "2020 ACS Estimates are derived from 5-Year ACS due to pandemic collection issues."),
          zeroGrob(),
          cpal_logo_text(), 
          ncol = 1,
          heights = c(1, 3, 30, 1, 1, 1, 2))


```
